---
title: "CalCurCEAS_drift-specific_Pm_take2"
author: "Kourtney Burger"
date: "2025-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code source

Replicated code from Jenny Trickey's repo [here](https://github.com/jtrickey86/SpermWhale_DASBR/blob/main/HICEAS_2023/HICEAS23_drift-specific_Pm.Rmd)

# Load Libraries and settings
```{r}
#Load libraries
library(PAMpal)
library(PAMmisc)
library(dplyr)
library(manipulate)

#Set time zone to UTC
Sys.setenv(TZ = 'UTC')
```

# Set global variables
Follow this folder structure for raw data
- Root directory
  - Binaries/
    - CalCurCEAS_###/
      - YYYYMMDD
        - PG Binary files
      - ...
  - Databases/
    - CalCurCEAS_###_2024_SpermWhale_PAM20214.sqlite3
    - ...
  - GPS Data/
    - CalCurCEAS_###_GPS.csv
    - ...
  - Pm_AcousticStudies/ (empty folder for code results to be saved to)
    
```{r}
# Path to binaries, databases, and GPS files
Binaries <- 'Z:/drifting_recorder/1_Binaries-Database/2024_CalCurCEAS/'
Databases <- 'C:/Users/pam_user/Desktop/CalCurCEAS_023 PG Rerun/'
GPS <- 'C:/Users/pam_user/Desktop/CalCurCEAS_023 PG Rerun/'

# Path to save acoustic studies to
save_path <- 'C:/Users/pam_user/Desktop/CalCurCEAS/acoustic studies'

# Global variable to drift ID, edit with each run
DriftID <- '023'
```


# Add GPS into databases
```{r}
#Copy and paste the below lines for each drift
DB <- paste0(Databases,'CalCurCEAS_', DriftID, '_2024_SpermWhale_PAM20214.sqlite3') #path to first deployment database
Track <- paste0(GPS, 'CalCurCEAS_', DriftID, '_GPS.csv') #path to first deployment GPS csv

addPgGps(db = DB, gps = Track, source ='csv')
```

# Run PAMpal

```{r}
#Create a settings object (loading databases and binaries) #current filter setting is for Pm data
myPpsL <- PAMpalSettings(sr_hz=384000, winLen_sec=.0025,
                         filterfrom_khz=2, filterto_khz=NULL)

#Extracting data to build the AcousticStudy
AS_data <- processPgDetections(myPpsL, mode='recording', id=paste0('CalCurCEAS_2024_', DriftID)) 

saveRDS(AS_data, file=file.path(save_path, paste0('CalCurCEAS_2024_', DriftID, '.rdata')))
```

# Remove Click Detector

```{r}
AS_data_filtered <- filter(AS_data, detectorName == 'Click_Detector_1') 

saveRDS(AS_data_filtered, file=file.path(save_path, paste0('CalCurCEAS_2024_', DriftID, '_Filtered.rdata')))
```

# Reduce data to single Channels

```{r}
#separate data by channels
ch1AS_data_filtered <- filter(AS_data_filtered, Channel %in% 1)
ch2AS_data_filtered <- filter(AS_data_filtered, Channel %in% 2)

#save single channel data
saveRDS(ch1AS_data_filtered, file=file.path(save_path, paste0('CalCurCEAS_2024_', DriftID, '_ch1.rdata')))
saveRDS(ch2AS_data_filtered, file=file.path(save_path, paste0('CalCurCEAS_2024_', DriftID, '_ch2.rdata')))
```

# Load data if neccessary

```{r}
#reload data from anywhere: get Data
AS_data <- readRDS(file.choose())
```
