---
title: "CalCurCEAS_bw_results"
editor: visual
---

# CalCurCEAS Beaked Whale Results

This code generates summary figures and tables along with relevant text for the beaked whale detections from the 2024 NOAA CalCurCEAS survey.

## Load libraries

```{r}
#| warning: false
library(ggplot2)
library(dplyr)
library(purrr)
library(marmap)       
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggnewscale)
library(here)
library(lubridate)
library(readxl)
library(PAMpal)
library(ggrepel)
library(ggspatial)
```

## Define Settings

```{r}
# Worksheet containing deployment details
deployment_info_file<-'C:/Users/annes/Documents/Github/TethysSAEL/Deployment Worksheets/DeploymentDetails.xlsx'
  
# Folder containing all *_GPS.csv files
gps_dir <- here('_data','GPS','DASBRs')

# Path containing acoustic study of all beaked whale detections
AcousticStudy_File <-'F:/Beaked_whale_data/AcousticStudy/CombinedStudy_03_EventswRecordingsGPS.rds'

# Output folder for the map files
output_dir <- here('_figs','beaked_whales')

# Output filename (without extension — both .png and .pdf will be saved)
output_filename <- "CalCurCEAS_bw_"

# Map dimensions in inches
map_width  <- 6
map_height <- 10

# Bathymetry resolution in arc-minutes (1 = finest/slowest, 4 = coarser/faster)
# For a publication figure, 1 is recommended. Use 4 for quick previews.
bathy_resolution <- 1

# Depth contour levels to draw (metres, negative = below sea level)
contour_depths <- c(-100, -500, -1000, -5000)
```

## Import Drift Track Data

```{r}
deployment_info <- read_excel(deployment_info_file,sheet='deployDetails') %>%
 filter(Status=="Complete", Project=="CalCurCEAS")%>%
  mutate(Deployment = as.integer(DeploymentID),
    # Match deployment numbers to the DriftName values in the GPS files
    DriftName = sprintf("CalCurCEAS_%03d", DeploymentID),
    # Treat all deployment/recovery times as UTC
    # Recode "NA" text to real NA before converting to numeric,
    # then treat the numeric serial as a UTC datetime
    Deployment_Date = as.POSIXct(
      as.numeric(na_if(Deployment_Date, "NA")) * 86400,
      origin = '1899-12-30', tz = 'UTC'),
    Recovery_Date = as.POSIXct(
      as.numeric(na_if(Recovery_Date, "NA")) * 86400,
      origin = '1899-12-30', tz = 'UTC'),
    Deployment_Latitude = as.numeric(Deployment_Latitude),
    Deployment_Longitude = as.numeric(Deployment_Longitude),
    Recovery_Latitude = as.numeric(Recovery_Latitude),
    Recovery_Longitude = as.numeric(Recovery_Longitude)) %>%
  dplyr::select(
    DriftName,
    Deployment_Date,
    Deployment_Latitude,
    Deployment_Longitude,
    Recovery_Date,
    Recovery_Latitude,
    Recovery_Longitude)
```

## Import Beaked Whale Detection Data

```{r}
#Load in saved Acoustic Study with GPS data
data<-readRDS(AcousticStudy_File)
detections<-getClickData(data)

detections <- detections %>%
  mutate(species = recode(species,
                          BB = "BBWH",
                          Bb = "BBWH",
                          ZC = "GOBW",
                          BW37V="HUBW",
                          BW43 = "GTBW",
                          MS = "STBWH",
                          BW="UNME"))
#Remove detections of "BW?", "Oo" and "Other"
detections<-detections %>%
  filter(!species %in% c('BW?','?BW','Oo','Other'))

table(detections$species)

#Create DriftName variables
detections<-detections %>%
  mutate(DriftName = sub("\\..*$", "", eventId))

#Save the mean lat/lon for each event
detections_summary <- detections %>%
  group_by(eventId) %>%
  summarise(
    Longitude = mean(Longitude, na.rm = TRUE),
    Latitude  = mean(Latitude, na.rm = TRUE),
    UTC       = mean(UTC, na.rm = TRUE),
    species   = first(species),
    DriftName = unique(DriftName),
    .groups = "drop"
  )

detections_hourly <- detections_summary %>%
  mutate(hour_bin = floor_date(UTC, unit = "12 hour")) %>%
  group_by(species, DriftName,hour_bin) %>%
  summarise(
    Longitude = mean(Longitude, na.rm = TRUE),
    Latitude  = mean(Latitude, na.rm = TRUE),
    UTC       = hour_bin[1],
    n_events  = n(),
    .groups = "drop"
  ) %>%
  dplyr::select(species, Longitude, Latitude, UTC, n_events)

#Divide species into separate dataframes for plotting
BB_summary<-dplyr::filter(detections_hourly,species=="BBWH")
ZC_summary<-dplyr::filter(detections_hourly,species=="GOBW")
MC_MS_MG_summary<-dplyr::filter(detections_hourly,species %in% c("HUBW","STBWH","GTBW"))

```

## Import Drift Track Data

```{r}
# Load all GPS tracks
gps_files <- list.files(gps_dir, pattern = "_GPS\\.csv$",
                        full.names = TRUE, ignore.case = TRUE)

gps_all <- map_dfr(gps_files, function(f) {
  df <- read.csv(f, stringsAsFactors = FALSE, colClasses = c(DeviceId = "character"))
  df$utc_parsed <- as.POSIXct(df$UTC, tz = "UTC", format = "%Y-%m-%d %H:%M:%S")
  df
}) %>%
  filter(!is.na(Latitude), !is.na(Longitude)) %>%
  arrange(DriftName, utc_parsed)


#######################
##Restrict to times that instruments were deployed 
gps_all <- gps_all %>%
  inner_join(
    deployment_info %>%
      dplyr::select(DriftName, Deployment_Date, Recovery_Date),
    by = "DriftName"
  ) %>%
  filter(utc_parsed >= Deployment_Date, utc_parsed <= Recovery_Date) %>%
  dplyr::select(-Deployment_Date, -Recovery_Date)

```

## **Define Map Settings**

```{r}
# ── COMPUTE MAP EXTENT ────────────────────────────────────────────────────────
# Auto-expand bounding box to include all GPS and detection points
all_lons <- c(gps_all$Longitude)
all_lats <- c(gps_all$Latitude)

lon_pad <- diff(range(all_lons, na.rm = TRUE)) * 0.1
lat_pad <- diff(range(all_lats, na.rm = TRUE)) * 0.1

xlim <- range(all_lons, na.rm = TRUE) + c(-lon_pad, lon_pad)
ylim <- range(all_lats, na.rm = TRUE) + c(-lat_pad, lat_pad)

# ── BATHYMETRY ────────────────────────────────────────────────────────────────

message("Fetching bathymetry data (this may take a moment)...")

bathy_raw <- getNOAA.bathy(
  lon1 = xlim[1], lon2 = xlim[2],
  lat1 = ylim[1], lat2 = ylim[2],
  resolution = bathy_resolution,
  keep = TRUE       # cache locally so re-runs are fast
)

# Convert to data frame for ggplot
bathy_df <- fortify(bathy_raw) %>%
  rename(lon = x, lat = y, depth = z)

# ── COASTLINE ─────────────────────────────────────────────────────────────────

coastline <- ne_countries(scale = "medium", returnclass = "sf")
states    <- ne_states(country = c("United States of America", "Canada", "Mexico"),
                       returnclass = "sf")

# ── START / END MARKERS ───────────────────────────────────────────────────────

track_endpoints <- gps_all %>%
  group_by(DriftName, DeviceName) %>%
  summarise(
    start_lon = first(Longitude), start_lat = first(Latitude),
    end_lon   = last(Longitude),  end_lat   = last(Latitude),
    start_time = min(utc_parsed),
    end_time = max(utc_parsed),
    .groups = "drop"
  )

track_endpoints_single <- track_endpoints %>%
  group_by(DriftName) %>%
  summarise(
    # Take start position from whichever device started earliest
    start_lon  = start_lon[which.min(start_time)],
    start_lat  = start_lat[which.min(start_time)],
    start_time = min(start_time, na.rm = TRUE),
    
    # Take end position from whichever device ended latest
    end_lon  = end_lon[which.max(end_time)],
    end_lat  = end_lat[which.max(end_time)],
    end_time = max(end_time, na.rm = TRUE),
    
    .groups = "drop"
  )

# Reshape to long format so one geom handles both
endpoints_long <- bind_rows(
  track_endpoints_single %>% transmute(DriftName, lon = start_lon, lat = start_lat,
                                point_type = "Start (▲)"),
  track_endpoints_single %>% transmute(DriftName, lon = end_lon,   lat = end_lat,
                                point_type = "End (▽)")
)

# ── COLOUR PALETTE FOR SPECIES ────────────────────────────────────────────────

species_codes <- sort(unique(detections$species))

# Colorblind-friendly palette (expand automatically if > 8 species)
base_palette <- c(
  "#E69F00", # orange
  "#CC79A7", # reddish purple,
  "#F0E442", # yellow
  "#D55E00", # vermillion
  "#009E73"  # bluish green
)

myColors<-c("#22A884FF","#440154FF","#E69F00","#7AD151FF","#FDE725FF", "#414487FF")

species_colors <- setNames(
  # colorRampPalette(base_palette)(length(species_codes)),
  #viridis(length(species_codes)),
  myColors,
  species_codes
)

```

## Map Plots

## BaseMap

Adjust code below so that a base map is made, and then species-specific detections, legends, and titles are plotted for each species map

```{r}

p <- ggplot() +
  # --- Bathymetry fill ---
  geom_raster(data = bathy_df %>% filter(depth < 0),
              aes(x = lon, y = lat, fill = depth),
              interpolate = TRUE) +
  scale_fill_gradientn(
    name   = "Depth (m)",
    colours = c("#08306b", "#08519c", "#2171b5", "#4292c6",
                "#6baed6", "#9ecae1", "#c6dbef", "#deebf7"),
    values  = scales::rescale(c(-6000, -4000, -2000, -1000, -500, -200, -50, 0)),
    limits  = c(min(bathy_df$depth, na.rm = TRUE), 0),
    guide   = guide_colorbar(barheight = 8, barwidth = 0.8,
                             title.position = "top",order=3)
  ) +
  
  # --- Bathymetric contours ---
  geom_contour(data = bathy_df %>% filter(depth < 0),
               aes(x = lon, y = lat, z = depth),
               breaks    = contour_depths,
               colour    = "white",
               linewidth = 0.2,
               alpha     = 0.5) +
  
  # --- Land (filled) ---
  geom_sf(data = coastline, fill = "grey75", colour = "grey40",
          linewidth = 0.3, inherit.aes = FALSE) +
  geom_sf(data = states,    fill = NA,       colour = "grey55",
          linewidth = 0.2, inherit.aes = FALSE) +
  new_scale_fill() +
  # --- Drift tracks ---
  geom_path(data = gps_all,
            aes(x = Longitude, y = Latitude, group = interaction(DriftName,DeviceName)),
            colour    = "white",
            linewidth = 0.6,
            alpha     = 0.8) +
  
  # --- Start / end triangles ---
  geom_point(data = endpoints_long,
             aes(x = lon, y = lat, shape = point_type),
             colour = "white",
             fill   = "black",
             size   = 2,
             stroke = 0.5) +
  scale_shape_manual(name   = "Track endpoint",
                     values = c("Start (▲)" = 24,   # filled up-triangle
                                "End (▽)"   = 22),
                     guide = guide_legend(order = 2)) + # filled down-triangle
  # ---Drift Number labels ---
  geom_text_repel(data = endpoints_long %>% filter(point_type == "Start (▲)") %>%
                    mutate(drift_num = as.integer(gsub(".*_(\\d+)$", "\\1", DriftName))),
                  aes(x = lon, y = lat, label = drift_num),
                  colour   = "white",
                  size     = 3,
                  max.overlaps = Inf,
                  box.padding=.5,
                  point.padding =.3,
                  fontface = "bold")+
  
  # --- Map extent ---
  coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
  
  # --- scale bar ---
  annotation_scale(
    location = "bl",  # bottom-left; also "br", "tl", "tr"
    width_hint=0.25,
    unit_category = "metric",
    style = "bar", # or "ticks"
    line_col = "white",
    text_col = "white",
    bar_cols = c("white", "black")
  )
  
```

### Goose-beaked whales

```{r}

plot_detections<-ZC_summary

FullMap_ZC <- p +
  
  # --- Detection points (colored by species) ---
  geom_point(data = plot_detections %>% filter(!is.na(Latitude)),
             aes(x = Longitude, y = Latitude,
                 fill = species),    # changed from colour to fill
             shape  = 21,
             colour = "black",              # fixed outline colour
             size   = 2,
             alpha  = 0.7) +
  scale_fill_manual(name   = "Species code",    # changed from scale_colour_manual
                    values = species_colors,
                    guide = guide_legend(order = 1)) +

  # --- Labels & theme ---
  labs(
    title    = "Drifting Recorder Tracks and Acoustic Detections",
    subtitle =  " Goose-beaked whale",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_bw(base_size = 11) +
  theme(
    panel.grid       = element_line(colour = "grey30", linewidth = 0.2,
                                    linetype = "dotted"),
    legend.position  = "right",
    legend.box       = "vertical",
    legend.key       = element_rect(fill = "grey20"),
    legend.background = element_rect(fill = "grey20", colour = NA),
    legend.text      = element_text(colour = "white"),
    legend.title     = element_text(colour = "white", face = "bold"),
    plot.background  = element_rect(fill = "white"),
    axis.text        = element_text(size = 8)
  )
```

### Baird's beaked whales

```{r}

plot_detections<-BB_summary

FullMap_BB <- p +
  
  # --- Detection points (colored by species) ---
  geom_point(data = plot_detections %>% filter(!is.na(Latitude)),
             aes(x = Longitude, y = Latitude,
                 fill = species),    # changed from colour to fill
             shape  = 21,
             colour = "black",              # fixed outline colour
             size   = 2,
             alpha  = 0.7) +
  scale_fill_manual(name   = "Species code",    # changed from scale_colour_manual
                    values = species_colors,
                    guide = guide_legend(order = 1)) +
  
  
  # --- Labels & theme ---
  labs(
    title    = "Drifting Recorder Tracks and Acoustic Detections",
    subtitle =  "Baird's beaked whale",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_bw(base_size = 11) +
  theme(
    panel.grid       = element_line(colour = "grey30", linewidth = 0.2,
                                    linetype = "dotted"),
    legend.position  = "right",
    legend.box       = "vertical",
    legend.key       = element_rect(fill = "grey20"),
    legend.background = element_rect(fill = "grey20", colour = NA),
    legend.text      = element_text(colour = "white"),
    legend.title     = element_text(colour = "white", face = "bold"),
    plot.background  = element_rect(fill = "white"),
    axis.text        = element_text(size = 8)
  )
```

### Hubbs', Stejneger's, and Gingko-toothed beaked whales

```{r}

plot_detections<- MC_MS_MG_summary

FullMap_BW <- p +
  
  # --- Detection points (colored by species) ---
  geom_point(data = plot_detections %>% filter(!is.na(Latitude)),
             aes(x = Longitude, y = Latitude,
                 fill = species),    # changed from colour to fill
             shape  = 21,
             colour = "black",              # fixed outline colour
             size   = 2,
             alpha  = 0.7) +
  scale_fill_manual(name   = "Species code",    # changed from scale_colour_manual
                    values = species_colors,
                    guide = guide_legend(order = 1)) +
  
  
  # --- Labels & theme ---
  labs(
    title    = "Drifting Recorder Tracks and Acoustic Detections",
    subtitle =  " Hubbs', Stejneger's and gingko-toothed beaked whale",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_bw(base_size = 11) +
  theme(
    panel.grid       = element_line(colour = "grey30", linewidth = 0.2,
                                    linetype = "dotted"),
    legend.position  = "right",
    legend.box       = "vertical",
    legend.key       = element_rect(fill = "grey20"),
    legend.background = element_rect(fill = "grey20", colour = NA),
    legend.text      = element_text(colour = "white"),
    legend.title     = element_text(colour = "white", face = "bold"),
    plot.background  = element_rect(fill = "white"),
    axis.text        = element_text(size = 8)
  )
```

## Save map figures

```{r}

dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

#Goose-beaked whale
png_path <- file.path(output_dir, paste0(output_filename, "FullMap_ZC.png"))
pdf_path <- file.path(output_dir, paste0(output_filename, "FullMap_ZC.pdf"))

ggsave(png_path, plot = FullMap_ZC, width = map_width, height = map_height,
       dpi = 300, bg = "white")
ggsave(pdf_path, plot = FullMap_ZC, width = map_width, height = map_height,
device = cairo_pdf)

#Baird's beaked whale
png_path <- file.path(output_dir, paste0(output_filename, "FullMap_BB.png"))
pdf_path <- file.path(output_dir, paste0(output_filename, "FullMap_BB.pdf"))

ggsave(png_path, plot = FullMap_BB, width = map_width, height = map_height,
       dpi = 300, bg = "white")
ggsave(pdf_path, plot = FullMap_BB, width = map_width, height = map_height,
       device = cairo_pdf)

#Other beaked whales
png_path <- file.path(output_dir, paste0(output_filename, "FullMap_BW.png"))
pdf_path <- file.path(output_dir, paste0(output_filename, "FullMap_BW.pdf"))

ggsave(png_path, plot = FullMap_BW, width = map_width, height = map_height,
       dpi = 300, bg = "white")
ggsave(pdf_path, plot = FullMap_BW, width = map_width, height = map_height,
       device = cairo_pdf)



```
