---
title: "Copy_midnight_noon_wav_files"
author: "Anne Simonis"
date: "2025-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

This document identifies `.wav` files whose filenames contain a 12-digit timestamp in the format `yymmddHHMMSS` (e.g., `241122143602`, as in `7939.241122143602.wav`). The script recursively scans a specified source directory and its subdirectories, extracts and parses the embedded timestamp from each filename, and selects files recorded during the time windows **00:00–00:59** or **12:00–12:59**.

Matching files are **copied** into a separate output directory while preserving the original subdirectory structure to avoid filename collisions. The workflow supports a **dry-run mode**, which reports the files that would be copied without modifying the filesystem, allowing results to be reviewed safely before execution.

Detectors can then be run in Pamguard on only the noon and midnight files.

```{r}
library(data.table)

# ---- USER SETTINGS ----
source_dir <- "E:/CalCurCEAS_018/"
target_dir <- "E:/CalCurCEAS_018_noon-and-midnight-files/"
dry_run    <- FALSE  # TRUE = preview only, FALSE = actually copy files
tz_use <- "UTC"     # change to Sys.timezone() if timestamps are local time


# ---- FIND WAV FILES RECURSIVELY ----
paths <- list.files(source_dir, recursive = TRUE, full.names = TRUE, no.. = TRUE)
paths <- paths[tolower(tools::file_ext(paths)) == "wav"]

dt <- data.table(
  path = paths,
  base = basename(paths)
)

# ---- EXTRACT 12-DIGIT TOKEN (yymmddHHMMSS) ----
# Safer pattern: token must be right before .wav (allows things like 7939.241122143602.wav)
dt[, token := fifelse(
  grepl("\\d{12}\\.wav$", tolower(base)),
  sub(".*(\\d{12})\\.wav$", "\\1", tolower(base)),
  NA_character_
)]


# ---- PARSE TOKEN ----
dt[, datetime := as.POSIXct(token, format = "%y%m%d%H%M%S", tz = tz_use)]

# Report/skip anything that didn't parse
bad <- dt[is.na(datetime)]
if (nrow(bad) > 0) {
  cat("Skipping", nrow(bad), "wav files (missing/invalid 12-digit timestamp). Examples:\n")
  print(bad[1:min(.N, 10), .(path, base)])
  cat("\n")
}

dt <- dt[!is.na(datetime)]

# ---- FILTER WINDOWS: 00:00–00:59 OR 12:00–12:59 ----
dt[, `:=`(
  hour   = as.integer(format(datetime, "%H")),
  minute = as.integer(format(datetime, "%M"))
)]

match <- dt[(hour %in% c(0L, 12L)) & (minute <= 59L)]

cat("Matched", nrow(match), "wav files in 00:00–00:59 or 12:00–12:59.\n")

if (nrow(match) == 0) {
  cat("Nothing to copy.\n")
} else {
  # ---- BUILD DEST PATHS (preserve relative dirs) ----
  src_norm <- normalizePath(source_dir, winslash = "/", mustWork = FALSE)
  match[, path_norm := normalizePath(path, winslash = "/", mustWork = FALSE)]
  match[, rel := substring(path_norm, nchar(src_norm) + 2)]
  match[, dest := file.path(target_dir, rel)]

  if (dry_run) {
    cat("\nDRY RUN — would copy (showing up to 50):\n")
    print(match[1:min(.N, 50),
                .(source = path,
                  destination = dest,
                  timestamp = format(datetime, "%Y-%m-%d %H:%M:%S"))])
    cat("\nTotal:", nrow(match), "files would be copied.\n")
  } else {
    # Create destination directories
    for (d in unique(dirname(match$dest))) {
      if (!dir.exists(d)) dir.create(d, recursive = TRUE, showWarnings = FALSE)
    }

    # Copy safely: overwrite = FALSE
    ok <- mapply(file.copy, match$path, match$dest,
                 MoreArgs = list(overwrite = FALSE), USE.NAMES = FALSE)

    cat(sum(ok), "of", length(ok), "files copied to", target_dir, "\n")
  }
}
```
