---
title: "DASBR_Glider_Overlap"
author: "Kourtney Burger"
date: "2025-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Visual overlap
We can see in our maps that there is spatial overlap between CalCurCEAS 7-15 
(some more than others). Characterizing the overlap is more difficult, how much 
overlap is enough to analyze the probability of detecting a species, etc.

# Prep data
## Load drifter data
```{r}
# Load driftwatch functions here - note: line 39 of this code is commented out. That only works if full driftwatch directory is saved to this repo but it is not needed for this purpose
source(here('_code/driftwatch/DriftWatchFunctions.R'))

# Load GPS database, may need to download updated version from driftwatch google drive if new drifts aren't included
db <- here('_data/GPS/DASBRs/SPOTGPS_Logger.sqlite3')

#All CalCurCEAS data
DASBR_gps <- getDbDeployment(db)
DASBR_gps <- filter(DASBR_gps, grepl('CalCurCEAS_', DriftName))

#Subset gliders that were in the same general area and remove lost drifts
DASBR_gps <- getDbDeployment(db, drift = c('CalCurCEAS_006', 'CalCurCEAS_008', 'CalCurCEAS_009', 'CalCurCEAS_010', 'CalCurCEAS_011', 'CalCurCEAS_012', 'CalCurCEAS_013', 'CalCurCEAS_014', 'CalCurCEAS_015', 'CalCurCEAS_016'))

#Subset to only include DriftName, UTC, latitude, and longitude
DASBR_gps <- DASBR_gps %>% 
  select(UTC,Latitude,Longitude,DriftName)

# Change DriftName to general name to match gliders and ship
DASBR_gps <- DASBR_gps %>% rename(
  ID = DriftName
)
```


## Load glider data
```{r}
# Load 3 glider gps files
Glider_sg639 <- read.csv(here('_data/GPS/Gliders/sg639_CalCurCEAS_Sep24_simpleSurfaceGPS.csv'))
Glider_sg679 <- read.csv(here('_data/GPS/Gliders/sg679_CalCurCEAS_Sep24_simpleSurfaceGPS.csv'))
Glider_sg680 <- read.csv(here('_data/GPS/Gliders/sg680_CalCurCEAS_Aug24_simpleSurfaceGPS.csv'))

# Add ID column to each dataframe
Glider_sg639 <- Glider_sg639 %>% 
  mutate(ID = 'sg639')

Glider_sg679 <- Glider_sg679 %>% 
  mutate(ID = 'sg679')

Glider_sg680 <- Glider_sg680 %>% 
  mutate(ID = 'sg680')

# Combine into one dataframe
glider_GPS <- bind_rows(Glider_sg639, Glider_sg679, Glider_sg680)

# Only interested in GPS points at start of dive, remove end of dive points
glider_GPS <- glider_GPS %>% 
  filter(label == "start")

# correct datetime format
glider_GPS$dateTime_UTC <- as.POSIXct(glider_GPS$dateTime_UTC, format = "%d-%b-%Y %H:%M:%S", tz = "UTC")

# Subset to only include glider ID, datetime, latitude, and longitude
glider_GPS <- glider_GPS %>% 
  select(ID, dateTime_UTC,latitude,longitude)

# change column names
glider_GPS <- glider_GPS %>% rename(
  ID = ID,
  UTC = dateTime_UTC,
  Latitude = latitude,
  Longitude = longitude
)
```

## Combine into one dataframe
```{r}
AllGPS <- bind_rows(DASBR_gps, glider_GPS)
```

# Check for overlap
## Set allowed distances
```{r}
# Max temporal distance (in seconds)
MaxTime <- 600 # 10 minutes
# Max spatial distance (in meters)
MaxSpace <- 100 # meters
```


## Temporal
```{r}

```

## Spatial
```{r}

```

## Temporal & Spatial
```{r}

```



#### TESTING
```{r}
### MATRIX IDEA NOT WORKING 


```

