---
title: "Format_bw_PACM"
editor: visual
---

# Intro

All beaked whale acoustic detection data must be formatted for the NOAA Fisheries Passive Acoustic Reporting System (PARS; <https://passiveacoustics.fisheries.noaa.gov/pars>). All detection data will ultimately be shared on the Passive Acoustic Cetacean Map (PACM; <https://passiveacoustics.fisheries.noaa.gov/pacm/>).

PARS Data Submission Guide available here: <https://passiveacoustics.fisheries.noaa.gov/pars/guide/>

This document takes Pamguard database files (.sqlite3) and generates data files ready to submit to PARs.

```{r study-definition}

#load required packages
library(dplyr)
library(RSQLite)
library(here)
library(stringr)
library(tidyverse)

source(here('_code','functions','read_pg_events.R'))

#define folder that contains all Pamguard databases
DBList<-dir(path='F:/Beaked_whale_data/Databases',pattern='*.sqlite3',full.names = TRUE)

#define folder that contains all GPS data
baseEffort<-'C:/Users/annes/Documents/Github/CalCurCEAS-2024-Drifter-Analysis/_data/GPS/DASBRs'

#define folder to save the data
savedir<-'F:/Beaked_whale_data/PACM/'
dir.create(savedir)

#Detection Data Codes
DetTemplate<-read_csv('F:/Beaked_whale_data/PACM/templates/detectiondata.csv',col_types = cols(.default = col_double()))
DetTemplate$deployment_code<-as.character()
DetTemplate$analysis_organization_code<-'SWFSC'
DetTemplate$analysis_sound_source_codes[1]<-list(list("GOBW","BWBB","BWMC","BW43","BWMS","BWC","UNBW"))
DetTemplate$analysis_sample_rate_khz<-384
DetTemplate$analysis_min_frequency_khz<-10
DetTemplate$analysis_max_frequency_khz<-384/2
DetTemplate$analysis_processing_code<-'Post-Processed'
DetTemplate$analysis_protocol_reference<-'SWFSC beaked whale protocol' 
DetTemplate$detector_code<-'PAMGuard Odontocete Detector'
DetTemplate$detector_version<-'2.02.14'
DetTemplate$detection_start_datetime<-as.Date.POSIXct(DetTemplate$detection_start_datetime)
DetTemplate$detection_end_datetime<-as.Date.POSIXct(DetTemplate$detection_end_datetime)
DetTemplate$detection_sound_source_code<-as.character(DetTemplate$detection_sound_source_code)
DetTemplate$detection_call_type_code<-'OD_CLICK_FM'
DetTemplate$detection_result_code<-'DETECTED' #other options include POSSIBLY_DETECTED and NOT_DETECTED
DetTemplate$localization_method_code<-'2D_TDOA'

```

## Process Pamguard Databases

```{r process-PG-db}

#Read Event data from all databases
bwData <- purrr::map_dfr(DBList, read_pg_events)

#Convert timestamp columns & calculate duration
bwData <- bwData |>
  dplyr::mutate(
    detection_start_datetime = as.POSIXct(UTC, format="%Y-%m-%d %H:%M:%OS", tz = "UTC"),
    detection_end_datetime = as.POSIXct(EventEnd, format="%Y-%m-%d %H:%M:%OS", tz = "UTC"), 
    detection_effort_secs=round(as.numeric(as.duration(interval(detection_start_datetime,detection_end_datetime))))
  )

#Rename columns to expected names for PACM
bwData <- bwData %>%
  rename('detection_sound_source_code' = 'eventType') %>% 
  dplyr::select('deployment_code','detection_start_datetime','detection_end_datetime','detection_effort_secs','detection_sound_source_code')
  
#Re-code species names
bwData <- bwData %>%
  mutate(species = recode(species,
                          "BB" = "BWBB",
                          "ZC" = "GOBW",
                          "BW39V"="BWMC",
                          "BW37V"="BWMC",
                          "BW43" = "BW43",
                          "MS" = "BWMS"
  ))
DetTemplate<-rbind(DetTemplate,bwData)

#Save Detections 
write.csv(bwData,file=here('_data','BeakedWhales','PACM','Detections.csv'))
```

Questions:

1.  What is the appropriate DetTemplate\$analysis_protocol_reference? (Simonis et al 2020?)
2.  Why are there two Stejneger's beaked whale codes? BWMS and STBWH
3.  Confirm: In Detections sheet, Analysis_start_datetime and analysis_end_datetime will be repeating values for each drift (i.e. All Drift 1 will have the same analysis start and end times)
